\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Hyperbolicity and the Green-Griffiths-Lang Conjecture}

%% All commands are in macros.sty

\section{The Kobayashi Pseudodistance}

\begin{defn}
A \textit{directed pair} $(X, V)$ is a pair of a complex manifold $X$ and a holomorphic subbundle $V \subset T_X$.
\end{defn}

Here let $\Delta$ be the unit disk in $\CC$ and $\rho$ the Poincare metric on $\Delta$.

\begin{defn}
Let $X$ be a complex manifold. The \textit{Kobayashi pseudodistance} is the pseudometric defined as follows:
\[ d_X(p,q) = \inf_\alpha \ell(\alpha) \]
where $\alpha$ is a chain of holomorphic disks $f_i : \Delta \to X$ and points $p = p_0, p_1, \dots, p_k = q$ of $X$ and pairs $a_1, b_1, \dots, a_k, b_k \in \Delta$ such that,
\[ f_i(a_i) = p_{i-1} \quad f_i(b_i) = p_i \]
and the length $\ell(\alpha)$ of the chain is defined as,
\[ \ell(\alpha) := \rho(a_1, b_1) + \cdots + \rho(a_k, b_k) \]
where $\rho$ is the Poincare metric on $\Delta$.
\end{defn}

\begin{example}
Let $X = \CC$ then $d_X = 0$. Indeed, by choosing larger and larger discs containing $p,q$ their pullback to the unit disk is then closer and closer to the origin and hence has vanishing Poincare distance.
\end{example}

\begin{rmk}
Recall the Schwarz--Pick lemma says that any holomorphic map $f : \Delta \to \Delta$ is a contraction for the Poincare metric. Therefore, $d_{\Delta} = \rho$.
\end{rmk}

\begin{lemma}
Let $f : X \to Y$ be holomorphic. Then $d_Y(f(x), f(y)) \le d_X(x,y)$
\end{lemma}

\begin{proof}
Indeed, choosing any chain of disks $g_i : \Delta \to X$ computing $d_X(x,y)$ we see that $f \circ g_i$ is a chain of disks connecting $f(x)$ and $f(y)$ of the same length. Therefore,
\[ d_Y(f(x), f(y)) = \inf_{\alpha} \ell(\alpha) \le d_X(x, y) \]
\end{proof}

\begin{cor}
If $f : \CC \to X$ is an entire curve then for $x, y \in f(\CC)$ we have $d_X(x,y) = 0$ meaning if $f$ is nonconstant then $d_X$ is degenerate along the image of $f$.
\end{cor}

\begin{proof}
Indeed, let $z_1, z_2 \in \CC$ map to $x,y$ respectively. Then,
\[ d_X(x,y) \le d_{\CC}(z_1, z_2) = 0 \]
\end{proof}

Brody's theorem is a converse to this result. We start by considering an infinitesimal analogue of the Kobayashi pseudodistance. Let $v \in T_{X, x_0}$ be a holomorphic tangent vector at $x_0 \in X$ and define,
\[ \kk_X(v) = \inf \{ \lambda > 0 \mid \exists f : \Delta \to X \text{ such that } f(0) = x_0 \text{ and } \lambda f'(0) = v \} \]
where $f : \Delta \to X$ is holomorphic. It is easy to check that holomorphic maps contract this pseudometric and for $X = \Delta$ it agrees with the Poincare metric.

\begin{theorem}
Let $X$ be a complex manifold. Then,
\[ d_X(p,q) = \inf_\gamma \int_\gamma \kk_X(\gamma'(t)) \d{t} \]
where the infimum is taken over all piecewise smooth curves joining $p$ and $q$.
\end{theorem}

\begin{defn}
A \textit{Brody curve} $f : \CC \to X$ is an entire curve which has bounded derivative with respect to some (equivalently, any) Hermitian metric.
\end{defn}

\begin{theorem}[Brody]
Let $X$ be a compact complex manifold. If $d_X$ is degenerate then there exists a Brody curve in $X$.
\end{theorem}

\begin{rmk}
Of course, in the case that $X$ is compact any entire curve is automatically Brody.
\end{rmk}



\section{Definitions}

\begin{defn}
We say a directed pair $(X, V)$ is,
\begin{enumerate}
\item \textit{Brody hyperbolic} if there does not exist a nonconstant entire map $f : \CC \to X$ tangent to $V$
\item \textit{Kobayashi hyperbolic} if the Kobayashi pseudodistance is nondegenerate (i.e. it is a metric).
\end{enumerate}
\end{defn}

\begin{theorem}[Brody]
Let $X$ be a compact complex manifold. Then $X$ is Kobayashi hyperbolic if and only if it is Brody hyperbolic.
\end{theorem}

\begin{rmk}
Therefore, we will call manifolds with this property just ``hyperbolic'' or ``analytically hyperbolic'' for emphasis.
\end{rmk}

\begin{defn}
If $X$ is a complex projective algebraic variety we say $(X, V)$ is
\begin{enumerate}
\item \textit{algebraically hyperbolic} if there exists $\epsilon > 0$ such that for every complete integral curve $C \subset X$ we have,
\[ 2 g(C) - 2 \ge \epsilon \deg_H{C} \]
where $g(C)$ is the geometric genus of $C$
\item \textit{algebraically quasi-hyperbolic} if $X$ contains finitely many genus $0$ and genus $1$ curves.
\end{enumerate}
\end{defn}

\begin{theorem}[Demailly]
Let $X$ be a smooth projective variety. Then the following hold,
\[ X \text{ is hyperbolic} \implies X \text{ is algebraically hyperbolic} \]
\end{theorem}

\begin{theorem}
If $X$ is algebraically hyperbolic then $X$ admits no nonconstant morphisms from an abelian variety.
\end{theorem}


Some references:
\begin{enumerate}
\item \chref{https://arxiv.org/pdf/math/0103084.pdf}{Xi Chen}
\item \chref{https://arxiv.org/pdf/1807.03665.pdf}{Javanpeykar}
\end{enumerate}

\subsection{The Green-Griffiths Locus and Jets}

\begin{theorem}[\chref{https://www-fourier.ujf-grenoble.fr/~demailly/manuscripts/hyperbolic.pdf}{Demailly's Notes} Theorem 7.9]
Let $(X, V)$ be a directed projective manifold and $A$ an ample line bundle. Then for any entire curve $f : \CC \to X$ tangent to $V$ and any $P \in H^0(X, E^{GG}_{k,m}(V^*) \ot A^{-1})$  we have $P(f', f'', \dots, f^{(k)}) = 0$ identically.
\end{theorem}

Therefore, if we fix an ample line bundle we can consider the locus cut out by all these differential equations.


\begin{defn}
The \textit{Green-Griffiths locus} $\GG_A(X, V)$ is the set $x \in X$ such that for all $k > 0$ there exists a $k$-jet $\varphi_k : (\CC, 0) \to (X, x)$ tangent to $V$ so that for all $m > 0$ every global jet differential $P \in H^0(X, E^{GG}_{k,m}(V^*) \ot A^{-1})$ satisfies $P(\varphi_k) = 0$.
\end{defn}

\begin{rmk}
The locus $\GG_A(X, V)$ is independent of the choice of ample line bundle (see \chref{https://arxiv.org/pdf/1302.4756.pdf}{Diverio and Rousseau} Lemma 2.2. This paper also gives many examples showing that $\Exc(X)$ can be strictly smaller than $\GG(X)$. However, it is conjectured that if $X$ is of general type then $\GG(X) \subsetneq X$.
\end{rmk}


LOOK AT THE HILBERT MODULAR SURFACES FOR WHICH THE GG LOCUS IS EVERYTHING


\section{Conjectures}

\begin{conj}[Kobayashi]
For $n \ge 2$ and $D \subset \P^n$ a very general hypersurface of degree $\deg{D} \ge 2n + 1$ then,
\begin{enumerate}
\item $D$ is hyperbolic
\item $\P^n \sm D$ is hyperbolic.
\end{enumerate}
\end{conj}

\begin{conj}[Green-Griffiths-Lang]
Let $X$ be a projective variety of general type. Then there exists a proper algebraic subvariety containing all non-constant entire curves $f : \CC \to X$.
\end{conj}

\begin{conj}[Demailly]
If $X$ is algebraically hyperbolic then $X$ is hyperbolic.
\end{conj}

\begin{prop}
The Green-Griffiths-Lang conjecture implies the Demailly conjecture.
\end{prop}

WHY?
\begin{proof}
Suppose $X$ is algebraically hyperbolic. If $X$ is not of general type then $X$ has a fibration over its canonical model by varities of Kodaira dimension $0$. (I NEED THAT IF NOT GENERAL TYPE THEN NOT ALGEBRAICALLY HYPERBOLIC DOES THIS FOLLOW FROM MMP)
\end{proof}

\section{Theorems}

\begin{theorem}[Bogomolov]
Let $X$ be a smooth projective surface with $s_2(X) = c_1(X)^2 - c_2(X) > 0$ then $X$ has finitely many genus $0$ or genus $1$ curves (i.e. it is algebraically quasi-hyperbolic).
\end{theorem}

\begin{theorem}[McQuillan]
Let $X$ be a smooth projective surface with $s_2(X) = c_1(X)^2 - c_2(X) > 0$ and $X$ has \textit{no} genus $0$ or genus $1$ curves then $X$ is hyperbolic.
\end{theorem}

Ingredients in Bogomolov's proof:
\begin{enumerate}
\item existence of a symmetric differential form $\omega \in H^0(X, S^m \Omega_X)$
\begin{enumerate}
\item stability of the cotangent bundle
\item Bogomolov vanishing theorem
\end{enumerate}
\item construction of foliations
\item Jouanolou's theorem on algebraic leaves of foliations
\end{enumerate}

\newpage

\section{Translation: Courbes de genre g\'{e}om\'{e}trique born\'{e} sur une surface de type g\'{e}n\'{e}ral}


The notion of stability of a point on a space of linear representations of a reductive group, due to Mumford [10], leads to a notion of stabilite for fiber bundles over a curve, whose properties were studied in [13] and [19].

\begin{defn}
Over a smooth proper integral curve, a vector bundle $E$ of rank $r(E)$ and degree $d(E)$ is \textit{stable} (resp. \textit{semistable}) if for every nonzero proper subbundle $F \subsetneq E$ we have,
\[ \frac{d(F)}{r(F)} < \frac{d(E)}{r(E)} \quad \left( \text{resp.} \frac{d(F)}{r(F)} \le \frac{d(E)}{r(E)} \right) \]
A vector bundle is \textit{unstable} if it is not semistable.
\end{defn}

Now let $X$ be a smooth proper surface over a field $k$, and $E$ a vector bundle over rank $2$ over $X$. Then a linear representation $\rho : \GL_2 \to \GL(V)$ produces an associated bundle $E^{(\rho)} := E \times_{\GL_2} V$ of rank $\dim{V}$.

\begin{defn}
We say a rank $2$ vector bundle is \textit{instable} if there exists a representation $\rho : \GL_2 \to \GL(V)$ with $\det{\rho} = 1$ such that $E^{(\rho)}$ admits a nonzero section which vanishes at some point.
\end{defn}

If the characteristic of $k$ is zero, which we will assume for the remainder, then Bogomolov's instabilite criterion is simply expressed in terms of devissage of bundles of rank $2$. It is interesting to note that we can here short-circuit the theory and prove directly using these simpler methods.
\par
There are many applications. We quote from memory a proof, elegant and algebraic, of the vanishing theorem of Kodaira-Ramanujan. In the remaining section we prove the following:

\begin{theorem}[0.3]
Let $X$ be a proper smooth surface of general type. Then $\Omega_X$ is not unstable.
\end{theorem}

As a consequence, we obtain the inequality $c_1^2 \le 4 c_2$ where $c_1, c_2$ are the Chern classes of the sheaf $\Omega_X^1$ -- improved by Miyaoka [9] which is the best form possible $c_1^2 \le 3 c_2$ - and a geometric result that we will develop here.
\par
The problem is the following: can we ``bound'' the family of curves of bounded geometric genus on a smooth proper surface $X$? We easily construct examples where the answer is negative. Bogomolov provides a partial solution in the case that $X$ is a surface of general type. We summarize briefly the method.
\par
Let $\pi : P = \P(\Omega_X^1) \to X$ be the canonical projection from the projectiviation of the canonical bundle. We construct on $P$ a good linear system of divisors allowing it to be mapped to the projective space $\P^N$. If $C$ is a smooth proper curve and $f: C \to X$ is a nonconstant morphism there is a lift $t_f : C \to P$ via the differential defined over points $\alpha \in P$ where $f$ is unramified as $t_f(\alpha) = (f(\alpha), f(v_\alpha))$ where $v_\alpha$ is a nonzero tangent vector to $C$ at $\alpha$. We apply this to the normalizations of curves embedded in $X$ and study their images in $\P^N$.
\bigskip\\
We prove the following result:

\begin{theorem}
Let $X$ be a smooth proper surface minimal of general type.
\begin{enumerate}
\item If $c_1^2 > c_2$ then the curves of bounded geometric genus on $X$ form a bounded family.
\item If $c_1^2 \le c_2$ and $\rank \NS{X} \ge 2$ then there exists a nonempty open cone $C \subset \NS{X}_{\RR}$ containing the cone $\{ z \mid z \in \NS{X}_{\RR} , z^2 \le 0 \}$ such that for any closed cone $C'$ contained in $C$ the family of curves of bounded geometric genus on $X$ have image in $\NS{X}_{\RR}$ contained in $C'$ forms a bounded family. Moreover, any translate of $C$ parallel to $K_X$ has the same property.
\end{enumerate}
\end{theorem}

As a corollary, we obtain finiteness of curves with negative self-intersection and bounded geometric genus on surfaces of general type. In particular a solution to Mordell's problem.
\par
Let's point out finally that Bogomolov uses a powerful result of Seidenberg on differential equations [18] but a recent paper of Jouanolou [5] allows us to avoid the use of this sledgehammer.

\subsection{Criteria for instability of vector bundles of rank $2$ on surfaces}

Considering the form of representations of $\PGL_2$ we give a definition equivalent to above.

\begin{defn}
A vector bundle $E$ of rank $2$ on a surface is \textit{unstable} if and only if there exists $n > 0$ such that $S^{2n} E \ot (\det{E})^{-n}$ has a nonzero section vanishing at some point of $X$.
\end{defn}

\subsubsection{Remark: devissage of vector bundles of rank $2$}

Let $E$ be a vector bundle of rank $2$ and $L$ an invertible sheaf and $s : L \to E$ a nonzero map. The bidual $M$ of $E/L$ is reflexive and hence invertible (since $X$ is a smooth surface), and the kernel $L_1$ of the homomorphism $E \to M$ is a larger invertible subsheaf of $E$ containing $L$. We say that it is a saturated line bundle of $E$. The cokernel $E/L_1$ is torsion-free in rank $1$, and hence of the form $I_Z \ot M$ for $M$ an invertible sheaf and $I_Z$ a sheaf of ideals defining a closed subscheme $Z$ of dimension $0$ outside of which $L_1$ is a subbundle of $E$. We have a diagram of exact sequences,
\begin{center}
\begin{tikzcd}
0 \arrow[r] & L \arrow[d] \arrow[r] & E \arrow[r] \arrow[d, equals] & E/L \arrow[d] \arrow[r] & 0
\\
0 \arrow[r] & L_1 \arrow[r] & E \arrow[r] & I_Z \ot M \arrow[r] & 0
\end{tikzcd}
\end{center}
We will say that the second line is a devissage of $E$. We can deduce the Chern classes of $E$,
\[ c_1(E) = c_1(L_1) + c_1(M) \quad \quad c_2(E) = c_1(L_1) \cdot c_1(M) + \deg{Z} \]

\begin{theorem}[Bogomolov-Mumford]
A vector bundle $E$ of rank $2$ over a surface $X$ is unstable if and only if there exists a devissage,
\[ 0 \to L \to E \to I_Z \ot M \to 0 \]
such that if $L' = L \ot M^{-1} = L^2 \ot (\det{E})^{-1}$ then either,
\begin{enumerate}
\item $L'$ is in the cone $C_+ \subset \NS{X}_{\Q}$ generated by positive divisors (the component of $C = \{ \alpha \mid \alpha^2 > 0\}$ containing the ample classes)
\item or $L' = \struct{X}$ and $Z$ is nonempty
\end{enumerate}
Moreover, the devissage is unique.
\end{theorem}

\begin{rmk}
A divisor class is \textit{positive} if and only if $D^2 > 0$ and $D \cdot H > 0$ for all ample $H$.
\end{rmk}

We will prove this using only Mumford's theory of instability.
\par
Let $P = \P(E)$ and $p : P \to X$ the projection and $\struct{P}(1)$ the canonical relatively ample bundle on $P$. A nonzero section $s \in H^0(X, S^{2n} E \ot (\det{E})^{-n})$ corresponds to a nonzero section $t \in H^0(P, \struct{P}(2n) \ot p^* (\det{E})^{-n})$. Let $\xi \in X$ be the generic point and $K = \kappa(\xi)$. If we choose a basis of $E_K$ then $s(\xi)$ corresponds to a homogeneous polynomial $F$ of degree $2n$ in two variables. Since $s$ is zero at some point of $X$, we know $s(\xi)$ is unstable for the action of $\PGL_2$ on $S^{2n} E_K \ot (\det{E_K})^{-n}$ (WHY?). We deduce from the stability criterion using 1-parameter subgroups [11] that $F$ has a root of order greater than $n$ in the algebraic closure of $K$, so also in $K$ (WHAT? WHY?), that's to say there exists an integer $r \ge 1$ and two polynomials $G, H$ homogeneous of degrees $1$ and $n-r$ respectively such that $F = G^{n+r} H$. Let $D$ be the divisor of $t$ and $\Delta$ the closure of the divisor defined over a generic point by $G$. We can write $D = (n + r) \Delta + \Delta'$ where $\Delta$ has degree $1$ and $\Delta'$ has degree $n - r$ on $P$. Therefore, there exist invertible modules $L, L'$ on $X$ such that,
\[ \struct{P}(\Delta) = \struct{P}(1) \ot p^* L \quad \struct{P}(\Delta') = \struct{P}(n-r) \ot p^* L' \]
and hence,
\[ (\det{E})^{-n} = L^{n+r} \ot L' \]
The divisor $\Delta$ corresponds to a section of $E \ot L$ and thus an injection $L^{-1} \embed E$ which by construction is saturated in $E$. We verify that it provides the desired devissage.


\subsection{Operations on unstable bundles}

Instability is preserved by passage to the dual and tensor product with a line bundle.

\begin{enumerate}
\item Let $f : Y \to X$ be a surjective morphism of surfaces, $E$ a vector bundle of rank $2$ over $X$. Then $E$ is unstable if and only if $f^* E$ is.

\item Let $f : Y \to X$ be a finite faithfully flat morphism of surfaces, $F$ a fiber bundle of rank $2$ on $Y$. Then if $F$ is unstable so is $f_* F$.
\end{enumerate}


\subsection{Proof of Theorem 0.3}

Suppose that $\Omega_X^1$ is unstable. Then there exists a devissage:
\[ 0 \to L \to \Omega^1_X \to I_Z \ot M \to 0 \]
and an integer $n > 0$ such that there is an injection $\cO_X \embed (L \ot M^{-1})^{\ot n}$. Note that $L \ot M^{-1}  = L^2 \ot (\det{\Omega^1_X})^{-1} = L^2 \ot (\Omega_X^2)^{\ot -1}$. Also, for $m \gg 0$,
\[ h^0(L^{2m}) = h^0((L \ot M^{-1})^{\ot m} \ot (\Omega^2_X)^{\ot m}) \ge h^0((\Omega_X^2)^{\ot m}) \in O(m^2) \]
Therefore, the theorem is a consequence of the following.

\begin{theorem}[Bogomolov]
Let $X$ be a smooth proper surface and $L \embed \Omega_X^1$ an invertible subsheaf. Then $h^0(L^n) \in O(n)$.
\end{theorem}

First recall the pretty result of Castelnuovo and of Franchis which we will need for the proof.

\begin{lemma}[4, 12]
Let $\omega_1, \omega_2$ be two holomorphic $1$-forms on $X$ which are linearly independent over $k$ such that $\omega_1 \wedge \omega_2 = 0$. Then there exists a curve $C$ which is proper and smooth over $k$ of genus $g \ge 2$ and two holomorphic $1$-forms $\theta_1, \theta_2$ on $C$ and a morphism $u : X \to C$ such that $\omega_i = u^* \theta_i$ for $i = 1,2$.
\end{lemma}

There exists a meromorphic function $f : X \rat \P^1$ such that $\omega_2 = f \omega_1$. This defines a morphism $f : X' \to \P^1$ where $X'$ is a blowup of $X$. Let $u : X' \to C \to \P^1$ be the Stein factorization.  We have an exact sequence of modules of differentials,
\[ 0 \to u^* \Omega^1_C \to \Omega^1_{X'} \to \Omega^1_{X'/C} \to 0 \]
We know $\omega_2 = f \omega_1$ and $0 = \d{\omega_2} = \d{f} \wedge \omega_1$ (since $\omega_i$ are global holomorphic forms they are closed by Hodge theory).

{\color{red} WHY DOES IT WORK ON AN OPEN}

But $\d{f}$ is pulled back from an open of $U$ so $\omega_1$ is also as it is parallel to $\d{f}$ hence also $\omega_2 = f \omega_1$. So above an open $U \subset C$ the forms $\omega_1, \omega_2$ are in the image of,
\[ H^0(u^{-1}(U), u^* \Omega_C^1) = H^0(U, \Omega^1_C) \to H^0(u^{-1}(U), \Omega^1_{X'}) \]
so we choose $\theta_1, \theta_2$ holomorphic forms on $U$ which pull back to $\omega_1, \omega_2$. However, $u_* \cO_{X'} = \cO_{C}$ so $\theta_1, \theta_2$ extend to global sections of $\Omega_C$ because $\omega_1, \omega_2$ are global sections of $\Omega_{X'}$. Indeed,
{\color{red} (WHY DOES IT EXTEND??) THIS SEEMS WRONG}


Since $\omega_1, \omega_2$ are $k$-independent so are $\theta_1, \theta_2$. Hence $g(C) \ge 2$ and therefore the map $u : X' \to C$ contracts all rational curves and hence factors through $X' \to X$ giving the required map.




\subsubsection{Completion of the Theorem}

Either, for all $n > 0$ we have $h^0(X, L^{\ot n}) \le 1$ or there exists $n > 0$ such that $h^0(X, L^{\ot n}) \ge 2$. In the second case, there is a standard method of extracting an $n^{\text{th}}$-root {\color{red} (WHAT THE HELL DOES THIS MEAN)} to get $h^0(X, L) \ge 2$. In this case, there are two forms $\omega_1, \omega_2 \in H^0(X, \Omega^1_X)$ such that $\omega_1 \wedge \omega_2 = 0$ since they arise from the same subsheaf of rank $1$. Therefore, by the lemma, there exists a curve $C$ and a morphism $u : X \to C$ and an invertible sheaf $L_0$ on $C$ such that,
\[ L \subset u^*(L_0) \]
{\color{red} AGAIN WHY?}
so we can conclude that,
\[ h^0(X, L^n) \le h^0(C, L_0^n) \in O(n) \]

\begin{cor}
If $c_1$ and $c_2$ are the Chern classes of $\Omega^1_X$ then $c_1^2 \le 4 c_2$.
\end{cor}

\subsubsection{Curves of bounded genus on a minimal surface of general type}

We provide a few examples showing that $X$ being general type plays an essential role, and that in the contrary case, there can be unbounded families of curves of fixed geometric genus.

\begin{example}
Let $X = \P^2$ then $\NS{X} = \Z$. There exist in the projective plane curves of bounded geometric genus but arbitrarily large degree.
\end{example}

\begin{example}
Let $E$ be an elliptic curve without complex multiplication and let $X = E \times E$. Then $\NS{X} = \Z f_1 \oplus \Z f_2 \oplus \Z \Delta$ where $f_i$ are the fiber classes and $\Delta$ is the diagonal. For every pair of integers $(m, n)$ the image in $X$ of the morphism $f_{m,n} : E \to X$ given by $f_{m,n}(\alpha) = (m \alpha, n \alpha)$ is a curve of class $m^2 f_1 + n^2 f_2 + (m-n)^2 \Delta$ and genus $1$.
\end{example}

\begin{example}
Let $B$ be a smooth proper curve and $\pi : X \to B$ a nonisotrivial (that is to say it does not become trivial after some finite base change $B' \to B$) minimal elliptic fibration admitting a section $\sigma : B \to X$ of infinite order. Let $\omega$ be the conormal bundle of $\sigma$. Then there exist global sections $g_2 \in H^0(X, \omega^4)$ and $g_3 \in H^0(X, \omega^6)$ such that $X$ is the minimal resolution of the surface $Y \subset \P_B(\omega^2 \oplus \omega^3 \oplus \cO_B)$ defined by the Weierstrass equation,
\[ y^2 z = x^3 - g_2 x z^2 - g_3 z^3 \]
Moreover, $\omega$ is independent of the section $\sigma$ and has degree $- \sigma(B)^2$. If the degree is zero, then $g_2$ and $g_3$ are constant and the fibration $\pi$ is isotrivial. There exist infinitely many sections of negative self-intersection and the classes are algebraically distinct.
\end{example}

Notation: we write $K_X$ for a canonical divisor of $X$ and $T_X$ the tangent bundle and $\pi : \P(\Omega_X) \to X$ the canonical projection and $L = \cO_P(1)$ the relatively ample bundle for $\pi$.
\par
Let $F$ be an invertible bundle on $X$. We note that $F$ is a divisor of some linear system {\color{red} WHAT'S THE CONTENT?}. Moreover, for any rational number $\ell \in \QQ$, we allow ourselves to form the sheaf $\ell F$, extending that we consider the tensor powers $(\ell F)^{\ot m}$ for which $m$ is such that $m \ell$ is an integer.


\subsection{Construction of a good linear system of divisors on $P$}

\begin{prop}
Let $F$ be an invertible sheaf on $X$ and $\ell$ a rational positive number such that,
\begin{enumerate}
\item $K \cdot F \ge 0$
\item $(K + 2 \ell F)^2 > 0$
\item $c_1^2(\Omega_X \ot \ell F) - c_2(\Omega_X \ot \ell F) > 0$
\end{enumerate}
Then for $m \gg 0$ the linear system $(L \ot \pi^* (\ell F))^m$ defines a rational map $u_F : \P(\Omega_X^1) \rat \P^N$ birational onto its image.
\end{prop}

\begin{proof}
By the theorem of Iitaka [20], it suffices to show that for $m \gg 0$,
\[ h^0(P, (L \ot \pi^* (\ell F))^m) = h^0(X, S^m(\Omega_X \ot \ell F)) \ge O(m^3) \]
The Riemann-Roch formula for $E$ shows that,
\[ \chi(S^m E) = \frac{m (m+1)(m+2)}{24} (c_1^2(E) - 4 c_2(E)) + \frac{m+1}{2} \left[ \frac{m^2}{4} c_1^2(E) - \frac{m}{2} K_X \cdot c_1(E) \right]  + (m + 1) \chi(\cO_X) \]
and hence for $m \gg 0$,
\[ h^0(S^m(\Omega_X^1 \ot \ell F)) + h^2(S^m(\Omega_X^1 \ot \ell F)) \sim h^1(S^m(\Omega_X^1 \ot \ell F)) + \frac{m^3}{6} \left[ c_1^2(\Omega_X \ot \ell F) - c_2(\Omega_X^1 \ot \ell F) \right] \ge O(m^3) \]
By Serre duality, {\color{red} HOW DO I FIX THE DUAL AND $S^m$ IN POSITIVE CHAR}
\[ h^2(S^m(\Omega_X \ot \ell F)) = h^0(K \ot S^m(T_X \ot -\ell F)) \]
Choosing some divisors $D$ and $D'$ ample and smooth such that,
\[ \cO_X(-D') \subset K \subset \cO_X(D) \]
we find that,
\[ \bigg| h^0(K \ot S^m(T_X \ot -\ell F)) - h^0(S^m(T_X \ot -\ell F)) \bigg| \in O(m^2) \]
Therefore, we conclude by appealing to the following lemma.
\end{proof}

\begin{lemma}
For any $m > 0$ we have $H^0(S^m(T_X \ot - \ell F)) = 0$.
\end{lemma}

\begin{proof}
We showed that $T_X \ot -\ell F$ is not unstable. Hence, the only sections of $H^0(X, S^{2m}(T_X \ot -\ell F) \ot \det{(T_X \ot -\ell F)}^{-m})$ are nowhere vanishing. If we show for $m \gg 0$ that $H^0(\det{(T_X \ot - \ell F)}^{-m})$ has a nonzero section with a zero at some point $x \in X$ then its product with a section $H^0(S^{2m}(T_X \ot - \ell F))$ will give a contradiction.
Thus, the result will follow from the definition,
% REVIEW: Should the exponent $-m$ be outside the \det? i.e. $\det(T_X \otimes -\ell F)^{-m} = \cO_X(m(K + 2\ell F))$? Also the RHS is a divisor class, not a line bundle -- consider using $\cO_X(m(K + 2\ell F))$.
\[ \det{(T_X \ot - \ell F)}^{-m} = m (K + 2 \ell F) \]
and Riemann-Roch,
\[ \chi(m(K + 2 \ell F)) \sim \frac{m^2}{2} (K + 2 \ell F)^2 \in O(m^2) \]
and therefore,
\[ h^0(m(K + 2 \ell F)) + h^2(m(K + 2 \ell F)) \ge O(m^2) \]
by Serre duality,
\[ h^2(m(K + 2 \ell F)) = h^0(K - m(K + 2 \ell F)) \]
Since $K \cdot (K - m(K + 2 \ell F)) = K^2 - m K \cdot (K + 2 \ell F) < 0$ and $K$ is nef (we assumed that $X$ is minimal) $h^0(K - m (K + 2 \ell F)) = 0$ for $m \gg 0$ giving the result.
\end{proof}

For any bundle $F$ verifying the conditions of the proposition, we fix, once and for all, $m$ and $\ell$ and let $Z_F$ be the closed subset of $\P(\Omega_X)$ outside of which $u_F$ is defined.

\begin{defn}
Let $C$ be a curve embedded in $X$ and $f : \wt{C} \to C$ its normalization. If $t_f(\wt{C})$ is not contained in (resp. is contained in) $Z_F$, we say that $C$ is $F$-regular (resp. $F$-irregular).
\end{defn}

\subsection{Proof of Theorem 0.4}

\begin{prop}
Let $F$ be an invertible bundle, and $\ell$ a rational number satisfying the conditions of 3.3.1 (above). For some positive real number $A$, the curves $C$ of geometric genus $g$ on $X$, such that $F \cdot C \le A$ form a bounded family.
\end{prop}

\begin{proof}
Let $C$ be a $F$-regular curve and $\wt{C}$ the normalization. The rational map
\[ \wt{C} \xrightarrow{t_f} \P(\Omega_X^1) \rat \P^N \]
extends to a morphism defined by a linear series contained in the sections of
\[ t_f^*(L \ot \pi^* \ell F)^m \]
where $m$ was chosen before so that $\P(\Omega_X) \rat \P^N$ is birational onto its image. Denote by $C'$ the image curve.
\par
By construction, $t_f^* L$ is the image of the tangent map
\[ \d{f} : f^* \Omega_X^1 \to \Omega_{\wt{C}}^1 \]
hence
\[ \deg{C'} \le \deg{t_f^*(L \ot \pi^* \ell F)^m} \le m \deg{(\Omega^1_{\wt{C}} \ot \ell F|_{\wt{C}})} = m (2g - 2) + m \ell F \cdot \wt{C} \le m (2g - 2) + m \ell A \]
Therefore, the curves form a bounded family. We also need to consider the $F$-irregular curves which is done later.
\end{proof}

The following theorem which we will prove later

\begin{theorem}[Bogomolov]
The family of $F$-irregular curves is bounded.
\end{theorem}

finishes the proposition. This result, along with the preceding proposition, applied to the case $F = 0$ (so that the existence of $A$ is automatic), finishes the proof of the first assertion of the main theorem.

\subsubsection{Proof in the case $c_1^2 \le c_2$ and $\rank \NS(X) \ge 2$}

Consider the equalities
\[ c_1(\Omega_X^1 \ot \ell F) = c_1 + 2 \ell F \]
and
\[ c_2(\Omega_X^1 \ot \ell F) = c_2 + \ell K_X \cdot F + \ell^2 F^2 \]
The conditions of 3.3.1 result in
\[ K_X \cdot F \ge 0 \quad (K_X + 2 \ell )^2 > \frac{4 c_2 - c_1^2}{3} \ge c_1^2 \]
where we use $c_1^2 \le c_2$. Consider the plane in $\NS(X)_{\RR}$ spanned by the images of $F$ and $K_X$, let us take an orthonormal basis formed by a vector collinear with $K_X$ and some orthogonal vector, in which the intersection form is given by the matrix
\[ \begin{pmatrix}
1 & 0
\\
0 & -1
\end{pmatrix} \]
The points $2 \ell F$ which verify the above inequalities are the interior of the hatched area (see figure). Let $C_0$ be the open cone tangent to the positive asymptotes of the hyperbola and centered at the origin. The line bundles $F$ verifying the inequalities are those whose images in $\NS(X)_{\RR}$ are contained in $C_0$. Let $C_1$ be the dual closed cone, and $C$ the complement, which contains the cone
\[ \{ z \in \NS(X)_{\RR} \mid z^2 \le 0 \} \]
where
\[ C_1 = \bigcap_{F \in C_0} \{ z \mid z \cdot F \ge 0 \} \]
and
\[ C = \bigcup_{F \in C_0} \{ z \mid z \cdot F < 0 \} \]
We finish the argument by compactness.
\par
The translates of $C$ parallel to $K_X$ are obtained by varying the constant $A$.

\begin{rmk}
Suppose $c_1^2 \le c_2$ and $\rank{\NS(X)} = 1$. Let $F$ be a numerically nontrivial line bundle over $X$. Then $F$ is numerically equivalent to $\lambda K_X$ for $\lambda \neq 0$. The condition $K_X \cdot G \ge 0$ implies $\lambda > 0$. Now it is well known that the curves verifying $K_X \cdot C \le A$ form a bounded family.
\end{rmk}

\begin{cor}
Let $X$ be a surface of general type such that $c_1^2 > c_2$. Then there exist only finitely many curves on $X$ of geometric genus $0$ or $1$.
\end{cor}

It is a consequence of the fact that a surface of general type is not ruled or elliptic [3], [17].

\begin{cor}
On a minimal surface of general type, a family of curves of bounded self intersection and bounded geometric genus is bounded.
\end{cor}

\begin{cor}[Problem of Mordell]
Let $X$ be a minimal surface of general type, $C$ a proper smooth curve and $f : X \to C$ a morphism with connected reduced fibers. Then if the fibration $f$ is nontrivial, there is a finite number of sections.
\end{cor}

\begin{proof}
Let $E$ be a section and $L = \struct{X}(E) \ot f^* \struct{E}(E)$. If $r = E^2 = \deg{\struct{E}(E)} > 0$ we verify that $L$ is numerically positive, so by the vanishing theorem of Kodaira-Ramanujam [15], $H^1(X, L^{-1}) = 0$. Furthermore, the Leray spectral sequence establishes
\[ H^1(X, L^{-1}) \iso H^0(C, R^1 f_* L^{-1}) \]
The exact sequence
\[ 0 \to \struct{X}(-E) \to \struct{X} \to \struct{E} \to 0 \]
gives $R^1 f_* \struct{X}(-E) = R^1 f_* \struct{X}$ hence
\[ R^1 f_* L^{-1} = \struct{E}(-E) \ot R^1 f_* \struct{X}(-E) \iso \struct{E}(-E) \ot R^1 f_* \struct{X} \]
From the exact sequence
\[ 0 \to \struct{X} \to \struct{X}(E) \to \struct{E}(E) \to 0 \]
we deduce the sequence
\[ 0 \to f_* \struct{X} \to f_* \struct{X}(E) \to \struct{E}(E) \to R^1 f_* \struct{X} \]
and since the genus of the fibers is nonzero, the map $\struct{E}(E) \to R^1 f_* \struct{X}$ is not zero, which contradicts the vanishing of $H^1(X, L^{-1})$.
\par
Therefore $E^2 \le 0$ (for another proof see [1] and [14]), if there are infinitely many they are contained in the fibers of some foliation on $X$ and then we check that $f$ must be a trivial fibration.
\end{proof}

\begin{rmk}
In characteristic $p$, M. Raynaud [16] used this procedure to construct a counter-example to the vanishing theorem. More precisely, he constructed a fibration $f : X \to C$ whose fibers are integral curves of nonzero arithmetic genus with a section $E$ such that $E^2 > 0$.
\end{rmk}

\begin{rmk}
If we do not bound the genus, the families of curves of negative self-intersection on a surface of general type can be unbounded: let $Y \to B$ be a nonisotrivial elliptic fibration having infinitely many section of negative self-intersection (see 3.1.3). There exists a cover $X \to Y$ ramified along a smooth hyperplane section which is a surface of general type. % REVIEW: Missing target of morphism "$Y \to$" -- should this be "$Y \to B$"? Also "finitely many" seems to contradict the point about unboundedness; should this be "infinitely many"?
The preimages of the sections of $Y \to B$ are curves such that at least one component has negative self-intersection, and there are infinitely many such curves on $X$.
\end{rmk}

\subsection{Study of $F$-irregular Curves}

{\color{red} TO FINISH}

\begin{defn}
An algebraic foliation on $X$ is defined by an invertible subsheaf $L$ of the sheaf $\Omega_X^1$. It is called \textit{saturated} if $L$ is a saturated subsheaf of $\Omega_X^1$.
\end{defn}

\begin{defn}
Let $C$ be a curve on $X$ and $\wt{C}$ the normalization, let $L \subset \Omega_X^1$ be an algebraic foliation. We say that $C$ is an integral curve of $L$ if the composition
\[ L|_{\wt{C}} \to \Omega^1_X |_{\wt{C}} \to \Omega^1_{\wt{C}} \]
is zero. The integral curves of a foliation are also integral curves of the associated saturated foliation and so are the loci where the map $L \to \Omega_X^1$ is zero.
\end{defn}

\begin{prop}
There exists a finite number of smooth proper surfaces $Z_i$ dominating $X$, each equipped with an algebraic foliation, and such that the $F$-irregular curves of $X$, except for finitely many, are the images in $X$ of integral curves of these foliations.
\end{prop}

\begin{proof}
The irreducible component of the closed set $Z_F \subset \P(\Omega_X^1)$ are curves or surfaces. The curves $C$ such that $t_f(\wt{C})$ is contained in the some component that does not dominate $X$ are finite in number.
\par
Let $Z_i$ be the desingularization of the components that dominate $X$ and $\alpha_i : Z_i \to X$ the resulting dominant morphisms, and $C$ a curve which lifts to $Z_i$
\begin{center}
\begin{tikzcd}
Z_i \arrow[r] \arrow[rd, "\alpha_i"] & \P(\Omega_X^1) \arrow[d, "\pi"]
\\
\wt{C} \arrow[u, "g"] \arrow[r, "f"] & X
\end{tikzcd}
\end{center}
On $\P(\Omega_X^1)$, we have by construction an exact sequence
\[ 0 \to L' \to \pi^* \Omega_X^1 \to L \to 0 \]
where $L'$ is an invertible sheaf.
\par
The injection
\[ L'|_{Z_i} \to \alpha_i^* \Omega_X^1 \to \Omega^1_{Z_i} \]
defines an algebraic foliation and we easily verify that $g(\wt{C})$ is an integral curve.
\end{proof}

To finish the proof of Theorem 0.4, two methods are possible -- we can, as is done by Bogomolov, use the following analytic result

\begin{theorem}[Seidenberg]
Let $A$ and $B$ be two elements of the complete local ring $R = k[[x,y]]$. There exists a scheme $S$ obtained from $\Spec{R}$ by a finite sequence of blowups at closed points located above the closed point of $\Spec{R}$ such that the strict transforms on $S$ of the analytic branches of solutions of the differential equation $A \d{x} = B \d{y}$ are, except for possibly finitely many, are pairwise disjoint.
\end{theorem}

then we show that a family $(C_i)_{i \in I}$ of pairwise disjoint curves on a surface is bounded.

Otherwise, one can use the following more precise - and less powerful statement:


\begin{theorem}[Jouanolou]
Let $L \to \Omega_X^1$ be an algebraic foliation on a smooth projective variety $X$ over $\CC$. If there exist infinitely many integral hypersurfaces, they are contained in the fibers of a (rational) contraction.
\end{theorem}

We will just consider the proof in the case that $X$ is a surface and $\L = \struct{X}$ -- meaning the foliation is defined by a global holomorphic $1$-form $\omega$, in order to lighten the notation. We now show that if there are infinitely many integral curves, there are meromorphic functions $f$ and $\varphi$ on $X$ such that $\omega = f \d{\varphi}$. Therefore, $\varphi$ defines the foliation.
\par
Let $\K_X$ be the sheaf of meromorphic functions on $X$ and $\Div(X)$ (resp.\ $\Div^{\tau}(X)$) the group of divisors (resp.\ numerically trivial divisors) on $X$ and $\Pic(X)$ (resp.\ $\Pic^{\tau}(X)$) the Picard group (resp.\ the group of numerically trivial line bundles) on $X$.
\par
The map of analytic sheaves (on the euclidean topology)
\[ \d{\log} : \struct{X}^\times \to \Omega_X^1 \]
defines a connecting map
\[ \Pic(X) = H^1(X, \struct{X}^\times) \to H^1(X, \Omega_X^1) \]

{\color{red} actually this works completely algebraically}

whose kernel contains $\Pic^{\tau}(X)$. There is a morphism of exact sequences:
\begin{center}
\begin{tikzcd}
0 \arrow[r] & \struct{X}^\times \arrow[r] \arrow[d, "\d{\log}"] & \K_X^\times \arrow[r] \arrow[d,"\d{\log}"] & \K_X^\times / \struct{X}^\times \arrow[r] \arrow[d] & 0
\\
0 \arrow[r] & \Omega_X^1 \arrow[r] & \Omega_X^1 \ot_{\struct{X}} \K_X \arrow[r] & \Omega_X^1 \ot (\K_X / \struct{X}) \arrow[r] & 0
\end{tikzcd}
\end{center}
Note that $H^0(X, \K_X^\times / \struct{X}^\times) = \Div(X)$ so the induced diagram of connecting maps
\begin{center}
\begin{tikzcd}
0 \arrow[r] & H^0(X, \struct{X}^\times) \arrow[r] \arrow[d] & H^0(X, \K_X^\times) \arrow[d] \arrow[r] & H^0(X, \K_X^\times / \struct{X}^\times) \arrow[r] \arrow[d] & \Pic(X) \arrow[d]
\\
0 \arrow[r] & H^0(X, \Omega_X^1) \arrow[r] & H^0(X, \Omega_X^1 \ot \K_X) \arrow[r] & H^0(X, \Omega_X \ot (\K_X / \struct{X})) \arrow[r] & H^1(X, \Omega_X^1)
\end{tikzcd}
\end{center}
produces a map
\[ \psi : \Div^{\tau}(X) \to H^0(X, \Omega_X^1 \ot \K_X) / H^0(X, \Omega_X^1) \]
by diagram chasing the kernel of
\[ H^0(X, \K_X^\times / \struct{X}^\times) \to H^1(X, \Omega_X^1) \]
must map to the image of
\[ H^0(X, \Omega_X^1 \ot \K_X) \to H^0(X, \Omega_X \ot (\K_X / \struct{X})) \]
which is then identified with the required quotient. $\psi$ satisfies the following properties:
\begin{enumerate}
\item if $D$ is an element of $\Div^{\tau}(X)$ and $\mu \in H^0(X, \Omega_X^1 \ot \K_X)$ is a representative of $\psi(D)$ then $\d{\mu} = 0$
\item if $D$ is defined on an open $U$ by the zeros of a function $g$ then $\d{\log{g}} - \mu$ is holomorphic on $U$
\item for any free subgroup $L \subset \Div^{\tau}(X)$ the morphism
% REVIEW: The subscript $G$ on $\psi_G$ seems inconsistent -- should this be $\psi|_L$ or $\psi_L$?
\[ \psi_L : L_{\CC} \to H^0(X, \Omega_X^1 \ot \K_X) / H^0(X, \Omega_X^1) \]
is injective.
\end{enumerate}

Let $M$ be the subgroup of $\Div(X)$ generated by the integral curves of $\F$ and $M^{\tau} := M \cap \Div^{\tau}(X)$. There is an exact sequence
\[ 0 \to M^{\tau} \to M \xrightarrow{c_1} H^2(X, \CC) \]
In the above, notation, for $D \in M^{\tau}$, the meromorphic $2$-form $\omega \wedge \mu$ can be locally written as
\[ \omega \wedge \d{\log}(g) - \omega \wedge v \]
where $v$ is holomorphic. This defines a morphism
\[ h : M^{\tau}_{\CC} \to H^0(X, \Omega_X^2) / \omega \wedge H^0(X, \Omega_X^1) \quad D \mapsto \omega \wedge v \]
To finish, it suffices to show

\begin{lemma}
If $\dim{\ker{h}} > 1$ then there exist meromorphic functions $f$ and $\varphi$ such that $\omega = f \d{\varphi}$.
\end{lemma}

\begin{proof}
Let $D_1$ and $D_2$ be linearly independent divisors in $\ker{h}$ and $\mu_1, \mu_2$ be the corresponding meromorphic $2$-forms. Since these live in the kernel, there exist holomorphic forms $\theta_i$ such that $\omega \wedge \mu_i = \omega \wedge \theta_i$ for $i = 1,2$ and hence nonzero meromorphic functions $k_i$ such that $\mu_i - \theta_i = k_i \omega$. Since $\d{\mu_i} = 0$ and $\d{\theta
_i} = 0$ (it is $\mu_i - \d{\log{g_i}}$ locally),
\[ \d(k_i \omega) = \d{k_i} \wedge \omega + k_i \d{\omega} = 0 \]
Thus,
\[ \d{k_i} \wedge \omega = - k_i \d{\omega} \]
and hence
\[ (k_1 \d{k_2} - k_2 \d{k_1}) \wedge \omega = 0 \]
which implies
\[ \omega \wedge \d{(k_2 / k_1)} = 0 \]
Let $\varphi = k_2 / k_1$. If $\varphi$ is constant then we deduce $\psi(D_2) = \varphi \psi(D_1)$ so $D_2 = \varphi D_1$ contradicting the independence. Hence $\varphi$ is nonconstant so $\omega \wedge \d{\varphi} = 0$ implies that $\omega = f \d{\varphi}$ for some meromorphic function $f$.
\end{proof}

This suffices because it proves that $M^{\tau}_{\CC}$ is finite dimensional unless $\omega$ arises from a rational map.

\section{Stability of the Cotangent Bundle}



% REVIEW: Section title "Miles Reid?" appears to be a placeholder -- should be given a proper title
\section{Miles Reid?}

\subsubsection{Interlude: regularizing meromorphic 1-forms via covers}

{\color{red} WHATIS THE CORRECT DEFINITION OF TAME?}

\begin{lemma}
Let $f : X \to Y$ be a morphism of locally noetherian schemes. If $Z \subset Y$ is an irreducible subset of codimension $\le r$ then either $f$ does not dominate $Z$ or there is some closed $Z' \subset X$ of codimension $\le r$ mapping to $Z$.
\end{lemma}

\begin{proof}
Using that $\codim{Z,Y} = \dim{\stalk{Y}{\xi}}$ where $\xi \in Z$ is the generic point we immediately reduce to the affine case. Either $\xi \notin f(X)$ and we are done or we can choose $f : U \to V$ a map of affine schemes sending $\xi' \in U$ to $\xi \in V$. Let $\varphi : A \to B$ be a map of noetherian rings and $\p \subset A$ a prime of height $\le r$ in the image of $\Spec{B} \to \Spec{A}$. Passing to $A_\p \to B_\p$ we need to find a prime $\q$ of $B_\p$ of height $\le r$ with $\varphi^{-1}(\q)$ maximal. Then $\p$ is the unique minimal prime over an ideal of definition $(x_1, \dots, x_r) \subset A_\p$ generated by at most $r$ elements by \chref{https://stacks.math.columbia.edu/tag/00KQ}{Tag 00KQ}. Since $B_\p / \p B_\p$ is nonzero (the fiber is nonempty) the ideal $(x_1, \dots, x_r) B_\p$ is proper hence, by the Krull height theorem, there exists a prime $\q$ containing it of height $\le r$. Then each $x_i \in \varphi^{-1}(\q)$ so $\p \subset \varphi^{-1}(\q)$ and we conclude.
\end{proof}

\begin{example}
Noetherianity is essential in the above. Indeed, we could take a domain $D$ with every nonzero prime of infinite height (as constructed in ``Anti-archimedean rings and power series rings'' by D.D. Anderson). Then for any nonzero nonunit $t \in D$ the map $k[t] \to D$ certainly falsifies the claim that the divisor $V(t)$ is in the image of a divisor (since there are none) although it is in the image of some prime.
\end{example}

\begin{prop}
Let $f : X \to Y$ be a proper dominant morphism of locally noetherian integral $S$-schemes that are smooth over $S$ at the generic points of all divisors. If $f$ is tame and $\omega \in (\Omega_{Y/S})_{\eta}$ is a meromorphic differential such that $f^* \omega \in H^0(X, \Omega_{X/S}^{\vee \vee})$ is a global reflexive differential then $\omega \in H^0(Y, \Omega_{Y/S}^{\vee \vee})$ is a global reflexive differential.
\end{prop}

\begin{proof}
Since $Y$ is regular in codimension $1$ it suffices to show that for each $\xi \in Y$ of height $1$ that $\omega_\xi \in (\Omega_Y)_{\xi}$. Since $f$ is proper and dominant it is surjective so we may choose $\xi' \in X$ mapping to $\xi$. The fiber over a divisor must contain a divisor of $X$ so we can choose $\xi'$ in the smooth locus, hence $f^* \omega$ is a well-defined differential form over $\stalk{X}{\xi'}$.  Since $\stalk{X}{\xi'}$ is a noetherian local domain by [Hartshorne, Ex.4.11] there exists a DVR $R \subset \Frac{\stalk{X}{\xi'}}$ dominating $\stalk{X}{\xi'}$.

{\color{red} FINISH}

\end{proof}

\begin{rmk}
For example, this holds for any tame dominant map of normal proper varieties over a perfect field.
\end{rmk}

{\color{red} COUNTEREXAMPLES}

\subsection{Proof of Theorem 0.4}

We suppose that $\L \embed \Omega_X^1$ is an invertible subsheaf. If $h^0(X, \L^{\ot n}) \le 1$ for all $n$ then we are done. Otherwise, there is some $n > 0$ such that $h^0(X, \L^{\ot n}) \ge 2$. In this case, by passing to a cyclic cover we may assume that $h^0(X, \L) \ge 2$. Therefore, there are two independent $1$-forms $\omega_1, \omega_2 \in H^0(X, \L) \subset H^0(X, \Omega^1_X)$ such that $\omega_1 \wedge \omega_2 = 0$ because they lie in the same $1$-dimensional subspace at the generic point $\L_{\eta} \subset \Omega_{X, \eta}^1$. Therefore, we may apply Castelnuovo's lemma to produce a morphism $f : X \to C$ to some curve of genus $g \ge 2$ with $\omega_1, \omega_2$ pulled back along $f$. By the proof of this lemma, we see that any local section of $\L$ is pulled back along $f$ hence $\L \embed f^* \Omega_C$.

\subsubsection{Ramified Cyclic Covers}

Let $X$ be a scheme and $\L \in \Pic{X}$ a line bundle and $s \in H^0(X, \L^{\ot n})$ a nonzero section of some tensor power. Then we may form a finitely-presented sheaf of $\struct{X}$-algebras,
\[ \cA = \struct{X} \oplus t \L^{\ot -1} \oplus \cdots \oplus t^{n-1} \L^{\ot -(n-1)} \]
where multiplication is defined in the obvious manner,
\[ (t^a f_1) (t^b f_2)  =
\begin{cases}
t^{a+b} f_1 f_2 & a + b < n
\\
% REVIEW: Since $0 \le a,b \le n-1$, we have $k \le 1$; the condition $nk \le a+b < (n+1)k$ for $k=1$ gives $n \le a+b < n+1$, which seems too restrictive. Should the second case simply be $a + b \ge n$ (with $k=1$)?
t^{a + b - nk} [(s^\vee)^{\ot k} \ot \id](f_1 f_2) & nk \le a + b < (n+1)k
\end{cases}  \]
where $[(s^\vee)^{\ot k} \ot \id] : \L^{\ot -(a+b)} \to \L^{\ot -(a + b - nk)}$. Then we define $X_{\L, s} := \rSpec{X}{\cA}$. Over the locus where $s$ is nonvanishing it is clear that $X_{\L, s} \to X$ is a degree $n$ cyclic cover which is \etale for $n$ nonzero in the base scheme.
\bigskip\\
Note that $\cA$ can also be described as follows. Consider the symmetric algebra,
\[ \Sym{\bullet}{\L^\vee} = \bigoplus_{n = 0}^\infty t^n \L^{\ot -n} \onto \cA \]
which is the quotient as a sheaf of algebras by the ideal generated by $(t^n f - s^\vee(f))$ for local sections $f$ of $\L^{\ot -n}$. Therefore, $X_{\L, s} \embed \V_X(\L)$ is a closed subscheme of the total space of the line bundle $\L$ which can be described as the locus of points $(x, v)$ such that $v^n = s(x)$.
\bigskip\\
Note that under $\pi : \V_X(\L) \to X$ we get a canonical section $t \in H^0(\V_X(\L), \pi^* \L)$ and hence for $f : X_{\L, s} \to X$ there is a canonical section $t \in H^0(X_{\L, s}, f^* \L)$ such that $t^n = f^* s$.
\bigskip\\
Now suppose that $s_1, s_2 \in H^0(X, \L^{\ot n})$ are two independent sections. Then by passing to the iterated cyclic cover, $X' = (X_{\L, s_1})_{f^* \L, f^* s_2} \to X_{\L, s_1} \to X$ we get $\L' = f^* \L$ and two canonical sections $t_1, t_2 \in H^0(X', \L')$ such that $t_i^n = f^* s_i$ for $i = 1,2$.
\bigskip\\
Furthermore, suppose that $n$ is invertible on the base and there is an injection $\L \embed \Omega_X^1$. Then passing to the cyclic cover (which is generically \etale) we get $f^* \L \embed f^* \Omega_X^1 \embed \Omega_{X'}^1$ which is injective because it is at the generic point. Hence we reduce to the situation that $h^0(X, \L) \ge 2$.

\section{Green-Griffiths Locus}

Let $(X,V)$ be a directed manifold and $A$ an ample line bundle. Let $GG_A(X,V)$ be the intersection over all $k$ of the image under $J_k(V) \to X$ of the $A$-augmented stable base locus of $\cO_{J_k(V)}(1)$ i.e. the intersection of the zero locus of all sections of $\cO_{J_k}(m) \ot \pi^* A^{-1}$ over all $m$.  Let $\Exc(X,V)$ be the union of all entire curves in $X$ tangent to $V$.

\begin{theorem}[Green-Griffiths]
$\Exc(X,V) \subset GG_A(X,V)$.
\end{theorem}

\chref{https://arxiv.org/pdf/1302.4756}{here} they prove the following

\begin{theorem}
Let $W \subset V$ be a saturated coherent subsheaf so that
% REVIEW: $W^{[-m]}$ is unusual notation -- should this be $S^m W^*$ or similar? Check the reference for the intended meaning.
\[ H^0(X, W^{[-m]} \ot A^{-1}) = 0 \]
for all $m > 0$ then $GG(X,V) = X$.
\end{theorem}

\begin{proof}
We get surjections
\[ E^{GG}_{k,m} T_X^* \to E^{GG}_{k,m} V^* \to E_{k,m}^{GG} W^* \]
Using the standard filtration, the assumption implies
\[ H^0(X, E_{k,m}^{GG} W^*) = 0 \]
for all $k, m$. Therefore, any jet tangent to $W$ living over $x \in X$ contributes to $GG_A(X,V)$ so we conclude. There is a little subtlety making this work when $W$ is not a subbundle, but then we can just shrink to throw out the codimension $2$ subset where it is not a subbundle and make the same argument since sections of the jet bundles extend.
\end{proof}

\begin{rmk}
Note that there is also a $DS(X,V)$ Demailly-Semple locus using ``invariant jets'' meaning sections of the Demailly-Semple jets.
\end{rmk}

\begin{example}
In particular, if $X = X_1 \times X_2$ then $GG(X) = X$. Indeed, choose $T_{X_1} \subset T_{X_1 \times X_2}$ as the subsheaf which has no embedding of $A$ even after taking tensor powers since it is pulled back from $X_1$.
\end{example}

Furthermore, we can take quotients of products of curves. For example $X = C^{(2)} = (C \times C) / S_2$ with $g(C) \ge 3$ then $X$ is general type, hyperbolic, and has $GG(X) = X$.

\subsection{Symmetric Domains}

\begin{defn}
A bounded domain $\Omega \subset \CC^n$ is called a \textit{bounded symmetric domain} if for all $x \in \Omega$ there is an involutive ($\sigma_x^2 = \id$) biholomorphism $\sigma_x : \Omega \to \Omega$ for which $x$ is an isolated fixed point.
\end{defn}

Question: is aut transitive?

\begin{rmk}
If $X$ is uniformized (without elliptic points) by a bounded symmetric domain $\Omega$ then any entire map $f : \CC \to X$ lifts to $f : \CC \to \Omega$ and hence is constant by Liouville's theorem. Hence $X$ is hyperbolic.
\end{rmk}

\begin{theorem}
If $X$ is uniformized by the $n$-ball $\mathbb{B}^n$ then $X$ has ample cotangent bundle. In particular, $GG(X) = \empty$ using jets of order $1$.
\end{theorem}

\begin{proof}
The Bergman metric on $\mathbb{B}^n$ is invariant under biholomorphisms and hence descends to $X$. It has negative sectional Riemannian curvature and hence $\Omega_X$ is ample by Kodaira's theorem.
\end{proof}

\begin{defn}
We say that $X = \Omega / \Gamma$ is \textit{reducible} if there is a finite index subgroup $\Gamma_0 \subset \Gamma$ and a decomposition $\Omega \cong \Omega_1 \times \Omega_2$ so that $\Gamma_0 \cong \Gamma_1 \times \Gamma_2$ acting as a product.
\end{defn}

\begin{theorem}
Let $X$ be a complex projective manifold uniformized by a reducible bounded symmetric domain. Then $GG(X) = X$.
\end{theorem}

\ifSubfilesClassLoaded{%
  \bibliographystyle{alpha}
  \bibliography{refs}
}{}

\end{document}
